---
title: "EMDS: Week 11 (fitting real data)"
output: html_notebook
---

This notebook contains code for simulation of a temporal discounting task and parameter recovery analysis for a single participant.

```{r, include=FALSE}
rm(list=ls())
library("dplyr")
library("ggplot2")
library("ggpubr")
library("rstan")
options(mc.cores = 4)
rstan_options(auto_write = TRUE)
# library("bridgesampling")
# library("bayesplot")
```

First, let's load our data.

```{r}
resp <- read.csv('td-responses.csv')
resp <- resp[,-1]
resp <- resp %>% filter(participant_id > 0)
```

First, let's estimate a so-called pooled model.

```{r}
choices <- resp[, 'response']
tp <- resp[, 5:8]
```

We recompile the Stan model (assuming all data are from a single ('representative') participant).

```{r}
# Compile Stan model
mod <- stan_model(file = 'td-exp-indiv.stan')
```

Next, we prepare the data for Stan.

```{r}
# arrange data
dd_dat <- list(T = length(choices),
               choices = choices,
               tp = tp)
```

Next, we estimate the posteriors using sampling.

```{r}
### Fit model
fit <- sampling(mod, data = dd_dat, chains = 4, iter = 2000)
```

Let's examine the traceplots.

```{r}
traceplot(fit, pars = c("r", "beta"), inc_warmup = TRUE)
```

```{r}
plot(fit, pars = c("r", "beta"), show_density = TRUE, ci_level = 0.95, fill_color = "purple")
fit
```
Extract parameter estimates (means of posteriors).

```{r}
param_est_summary <- summary(fit, pars = c("r", "beta"))$summary
param_est_mean <- param_est_summary[, c("mean")]
param_est_mean
```

Now let's estimate a hierarchical model.

First, we need to arrange the data.

```{r}
pIDs <- unique(resp['participant_id'])
choices <- matrix(0, nrow = dim(pIDs)[1], ncol = 30)
Tpart <- rep(0, dim(pIDs)[1])
TP <- list()

for (i in 1:dim(pIDs)[1]) {
  p <- pIDs[i,1]
  p_data <- resp %>% filter(participant_id == p)
  nt <- min(30, dim(p_data)[1])
  Tpart[i] <- nt
  choices[i, 1:nt] <- p_data[1:nt, 'response'] 
  tp <- matrix(0, nrow = 30, ncol = 4)
  tp[1:nt, 1:4] <- as.matrix(p_data[1:nt, 5:8])
  TP[[i]] <- tp
}
```

```{r}
stan_mod <- stan_model(file = 'td-hierarch-model.stan')
  
sim_dat <- list(N = dim(pIDs)[1],
                T = 30,
                Tpart = Tpart,
                choice = choices,
                TP = TP)

fit03 <- sampling(stan_mod, data = sim_dat, chains = 4, iter = 5000)
```


```{r}
traceplot(fit03, pars = c("r", "beta"), inc_warmup = FALSE)

fit_summary <- summary(fit03)
fit_summary

plot(fit03, pars = c("r", "beta"), show_density = TRUE, ci_level = 0.95, fill_color = "purple")
plot(fit03, pars = c("r"), show_density = TRUE, ci_level = 0.95, fill_color = "purple")

mean(extract(fit03, pars = "beta")[[1]])
quantile(extract(fit03, pars = "r")[[1]], probs = c(0.1, 0.9))
```
